---
title: "Data Cleaning and EDA"
author: "Umi Yamaguchi, Benjamin Rodovinski"
output: pdf_document
---

```{r}
# Load the dataset
data <- read.csv("data_academic_performance.csv", header=T)

head(data)
str(data)
summary(data)

```

```{r}
data$GENDER <- as.factor(data$GENDER)
levels(data$GENDER) <- c(0, 1)  #0 for Male and 1 for Female

```

```{r}
library(dplyr)

data <- data %>%
  mutate(EDU_MOTHER = ifelse(EDU_MOTHER == "Ninguno", "Not sure", EDU_MOTHER),
         EDU_FATHER = ifelse(EDU_FATHER == "Ninguno", "Not sure", EDU_FATHER))


data$EDU_FATHER <- as.factor(data$EDU_FATHER)
data$EDU_MOTHER <- as.factor(data$EDU_MOTHER)
data$OCC_FATHER <- as.factor(data$OCC_FATHER)
data$OCC_MOTHER <- as.factor(data$OCC_MOTHER)




data$PEOPLE_HOUSE <- case_when(
  data$PEOPLE_HOUSE == "One" ~ 1,
  data$PEOPLE_HOUSE == "Two" ~ 2,
  data$PEOPLE_HOUSE == "Three" ~ 3,
  data$PEOPLE_HOUSE == "Four" ~ 4,
  data$PEOPLE_HOUSE == "Five" ~ 5,
  data$PEOPLE_HOUSE == "Six" ~ 6,
  data$PEOPLE_HOUSE == "Seven" ~ 7,
  data$PEOPLE_HOUSE == "Eight" ~ 8,
  data$PEOPLE_HOUSE == "Nueve" ~ 9,  # Spanish for "Nine"
  data$PEOPLE_HOUSE == "Ten" ~ 10,
  data$PEOPLE_HOUSE == "Once" ~ 11,  # Spanish for "Eleven"
  data$PEOPLE_HOUSE == "Twelve or more" ~ 12,
  data$PEOPLE_HOUSE == "0" ~ 0,       # Handling entry with "0"
  TRUE ~ NA_real_  # Assign NA to any unexpected values
)

# Ensure PEOPLE_HOUSE is numeric
data$PEOPLE_HOUSE <- as.numeric(data$PEOPLE_HOUSE)

```

```{r}
data <- data %>%
  mutate(STRATUM = as.factor(as.numeric(gsub("Stratum ", "", STRATUM))))
```

```{r}
data <- data %>%
  mutate(OVERALL_SCORE = QR_PRO + CR_PRO + CC_PRO + ENG_PRO + WC_PRO)
```

```{r}
data$COMPUTER <- ifelse(data$COMPUTER == "Yes", 1, 0)
data$CAR <- ifelse(data$CAR == "Yes", 1, 0)
data$MOBILE <- ifelse(data$MOBILE == "Yes", 1, 0)

```

```{r}
df <- data %>%
  dplyr::select(GENDER, EDU_FATHER, EDU_MOTHER, OCC_FATHER, OCC_MOTHER, STRATUM, PEOPLE_HOUSE, COMPUTER, CAR, MOBILE, UNIVERSITY, OVERALL_SCORE)

```


```{r}
sum(is.na(df))
write.csv(df, "data.csv")
```


# Residual vs Fitted

```{r}
model <- lm(OVERALL_SCORE ~ GENDER + EDU_FATHER + EDU_MOTHER + 
            STRATUM + PEOPLE_HOUSE + COMPUTER + MOBILE + CAR + UNIVERSITY + OCC_FATHER + OCC_MOTHER, data =df)
```


```{r}
y_value <- resid(model)
x_value <- fitted(model)

plot(x = x_value, y = y_value, main="Residual vs Fitted", xlab="Fitted",
     ylab="Residuals")
```

```{r}
plot(model, which = 3)
```


```{r}
par(mfrow=c(1,2))
qqnorm(resid(model))
qqline(resid(model))

hist(df$OVERALL_SCORE, xlab="Overall Score", main="Histogram of Score")
```

## Log Transformation
```{r}
df$log_OVERALL_SCORE <- log(df$OVERALL_SCORE)

model_log <- lm(log_OVERALL_SCORE ~ GENDER + EDU_FATHER + EDU_MOTHER + 
            STRATUM + PEOPLE_HOUSE + COMPUTER + MOBILE + CAR + UNIVERSITY + OCC_FATHER + OCC_MOTHER, data = df)

# Generate the QQ plot
qqnorm(resid(model_log), main = "QQ Plot of Residuals (Log Transformed)")
qqline(resid(model_log))

hist(df$log_OVERALL_SCORE, xlab = "Log(Overall Score)", main = "Histogram of Log(Overall Score)", 
     col = "lightblue", border = "black", breaks = 30)


y_value <- resid(model_log)
x_value <- fitted(model_log)

plot(x = x_value, y = y_value, main="Residual vs Fitted", xlab="Fitted",
     ylab="Residuals")

```




## BoxCox Transformation

```{r}
 model_cox <- lm(OVERALL_SCORE ~ GENDER + EDU_FATHER + EDU_MOTHER + 
            STRATUM + PEOPLE_HOUSE + COMPUTER + MOBILE + CAR + UNIVERSITY + OCC_FATHER + OCC_MOTHER, data =data)

 boxcox_result <- boxcox(model_cox, lambda = seq(-2, 2, by = 0.1))

 best_lambda <- boxcox_result$x[which.max(boxcox_result$y)]
 print(best_lambda)

 df$transformed_SCORE <- (data$log_OVERALL_SCORE^best_lambda - 1) / best_lambda

 model_boxcox <- lm(transformed_SCORE ~ GENDER + EDU_FATHER + EDU_MOTHER +
             STRATUM + PEOPLE_HOUSE + COMPUTER + CAR + MOBILE + UNIVERSITY + OCC_FATHER + OCC_MOTHER, data =data)
```
```{r}
plot(fitted(model_boxcox), resid(model_boxcox), 
     main="Residuals vs Fitted (Box-Cox Transformed)", 
     xlab="Fitted", ylab="Residuals")

qqnorm(resid(model_boxcox), main="QQ Plot of Residuals (Box-Cox Transformed)")
qqline(resid(model_boxcox))
```





## BoxCox Transformation with log Transformation
```{r}
 model_cox <- lm(log_OVERALL_SCORE ~ GENDER + EDU_FATHER + EDU_MOTHER + 
            STRATUM + PEOPLE_HOUSE + COMPUTER + MOBILE + CAR + UNIVERSITY + OCC_FATHER + OCC_MOTHER, data =data)

 boxcox_result <- boxcox(model_cox, lambda = seq(-2, 2, by = 0.1))

 best_lambda <- boxcox_result$x[which.max(boxcox_result$y)]
 print(best_lambda)

 df$transformed_SCORE <- (data$log_OVERALL_SCORE^best_lambda - 1) / best_lambda

 model_boxcox <- lm(transformed_SCORE ~ GENDER + EDU_FATHER + EDU_MOTHER +
             STRATUM + PEOPLE_HOUSE + COMPUTER + CAR + MOBILE + UNIVERSITY + OCC_FATHER + OCC_MOTHER, data =data)
```


```{r}
plot(fitted(model_log), resid(model_log), 
     main="Residuals vs Fitted (Log Transformed)", 
     xlab="Fitted", ylab="Residuals")

qqnorm(resid(model_log), main="QQ Plot of Residuals (Log Transformed)")
qqline(resid(model_log))
```
## Square root Transformation

```{r}
data$sqrt_OVERALL_SCORE <- sqrt(data$OVERALL_SCORE)

# Fit the model with the square root transformed response variable
model_sqrt <- lm(sqrt_OVERALL_SCORE ~ GENDER + EDU_FATHER + EDU_MOTHER + 
            STRATUM + PEOPLE_HOUSE + COMPUTER + MOBILE + CAR + UNIVERSITY + OCC_FATHER + OCC_MOTHER, data = data)

plot(fitted(model_sqrt), resid(model_sqrt), 
     main="Residuals vs Fitted (Sqrt Transformed)", 
     xlab="Fitted", ylab="Residuals")


qqnorm(resid(model_sqrt), main="QQ Plot of Residuals (Sqrt Transformed)")
qqline(resid(model_sqrt))


hist(data$sqrt_OVERALL_SCORE, xlab = "Sqrt(Overall Score)", main = "Histogram of Sqrt(Overall Score)", 
     col = "lightgreen", border = "black", breaks = 30)

```

```{r}
plot(model, which = 3)
```



```{r}
head(df)
str(df)
summary(df)
```

```{r}
# Install
install.packages("ggplot2")
```
```{r}
library("ggplot2")
```


## Bar plot for Stratum (economic and social status): 
```{r}
# Bar plot for Stratum
ggplot(df, aes(x = factor(STRATUM))) +
  geom_bar(fill = "red", color = "black") +
  ggtitle("Distribution of Stratum") +
  xlab("Stratum (1: Low Income, 6: High Income)") +
  ylab("Count")


```

## Overall Score by Stratum (economic and social status)
```{r}
# Boxplot for Overall Score by Stratum
ggplot(df, aes(x = factor(STRATUM), y = OVERALL_SCORE)) +
  geom_boxplot(fill = "orange", color = "black") +
  ggtitle("Overall Score by Stratum") +
  xlab("Stratum") +
  ylab("Overall Score")

```


## Overall Score distribution by Gender
```{r}
# Facet grid for Overall Score distribution by Gender
ggplot(df, aes(x = OVERALL_SCORE)) +
  geom_histogram(binwidth = 10, fill = "purple", color = "black") +
  ggtitle("Overall Score by Gender") +
  facet_grid(~ GENDER) +
  xlab("Overall Score") +
  ylab("Count")

```


## Father's Education vs Father's Occupation
```{r}
clean_data <- data[!is.na(data$EDU_FATHER) & !is.na(data$OCC_FATHER), ]

ggplot(clean_data, aes(x = OCC_FATHER, fill = EDU_FATHER)) +
  geom_bar(position = "dodge") +  
  labs(title = "Father's Education vs Father's Occupation",
       x = "Father's Occupation",
       y = "Count",
       fill = "Father's Education") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```


## Mother's Education vs Mother's Occupation
```{r}
clean_data <- data[!is.na(data$EDU_MOTHER) & !is.na(data$OCC_MOTHER), ]

ggplot(clean_data, aes(x = OCC_MOTHER, fill = EDU_MOTHER)) +
  geom_bar(position = "dodge") + 
  labs(title = "Mother's Education vs Mother's Occupation",
       x = "Mother's Occupation",
       y = "Count",
       fill = "Mother's Education") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```


## Overall Score by Father's Occupation
```{r}
clean_data <- df[!is.na(df$EDU_FATHER) & !is.na(df$OCC_FATHER) & 
                   !is.na(df$EDU_MOTHER) & !is.na(df$OCC_MOTHER) & 
                   !is.na(df$OVERALL_SCORE), ]

ggplot(clean_data, aes(x = OCC_FATHER, y = OVERALL_SCORE)) +
  geom_boxplot() +
  labs(title = "Overall Score by Father's Occupation",
       x = "Father's Occupation",
       y = "Overall Score") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))  # Rotate x-axis labels for readability



```


## Overall Score by Mother's Occupation
```{r}

ggplot(clean_data, aes(x = OCC_MOTHER, y = OVERALL_SCORE)) +
  geom_boxplot() +
  labs(title = "Overall Score by Mother's Occupation",
       x = "Mother's Occupation",
       y = "Overall Score") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))  # Rotate x-axis labels for readability


```


## Overall Score by Mother's Education
```{r}

ggplot(clean_data, aes(x = EDU_MOTHER, y = OVERALL_SCORE)) +
  geom_boxplot() +
  labs(title = "Overall Score by Mother's Education",
       x = "Mother's Education",
       y = "Overall Score") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))  # Rotate x-axis labels for readability

```



## Overall Score by Father's Education
```{r}

ggplot(clean_data, aes(x = EDU_FATHER, y = OVERALL_SCORE)) +
  geom_boxplot() +
  labs(title = "Overall Score by Father's Education",
       x = "Father's Education",
       y = "Overall Score") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))  # Rotate x-axis labels for readability


```

```{r}
par(mfrow=c(2,2))

ggplot(df, aes(x = factor(COMPUTER))) + 
  geom_bar(fill = "lightblue", color = "black") +
  ggtitle("Distribution of Computer Access") +
  xlab("Computer Access (0 = No, 1 = Yes)") +
  ylab("Count")

ggplot(df, aes(x = factor(CAR))) + 
  geom_bar(fill = "lightgreen", color = "black") +
  ggtitle("Distribution of Car Ownership") +
  xlab("Car Ownership (0 = No, 1 = Yes)") +
  ylab("Count")

ggplot(df, aes(x = factor(MOBILE))) + 
  geom_bar(fill = "lightpink", color = "black") +
  ggtitle("Distribution of Mobile Phone Ownership") +
  xlab("Mobile Phone Ownership (0 = No, 1 = Yes)") +
  ylab("Count")

ggplot(df, aes(x = factor(UNIVERSITY))) + 
  geom_bar(fill = "lightcoral", color = "black") +
  ggtitle("Distribution by University") +
  xlab("University") +
  ylab("Count")
```

```{r}
# Bar plot for STRATUM and EDU_MOTHER
ggplot(df, aes(x = factor(STRATUM), fill = EDU_MOTHER)) +
  geom_bar(position = "dodge") +
  ggtitle("Stratum vs Mother's Education") +
  xlab("Stratum") +
  ylab("Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Bar plot for STRATUM and EDU_FATHER
ggplot(df, aes(x = factor(STRATUM), fill = EDU_FATHER)) +
  geom_bar(position = "dodge") +
  ggtitle("Stratum vs Father's Education") +
  xlab("Stratum") +
  ylab("Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}

# Box plot for OVERALL_SCORE vs COMPUTER
ggplot(df, aes(x = factor(COMPUTER), y = OVERALL_SCORE)) +
  geom_boxplot(fill = "lightblue", color = "black") +
  ggtitle("Overall Score by Computer Access") +
  xlab("Computer Access (0 = No, 1 = Yes)") +
  ylab("Overall Score")



# Box plot for OVERALL_SCORE vs CAR
ggplot(df, aes(x = factor(CAR), y = OVERALL_SCORE)) +
  geom_boxplot(fill = "lightgreen", color = "black") +
  ggtitle("Overall Score by Car Ownership") +
  xlab("Car Ownership (0 = No, 1 = Yes)") +
  ylab("Overall Score")



# Box plot for OVERALL_SCORE vs MOBILE
ggplot(df, aes(x = factor(MOBILE), y = OVERALL_SCORE)) +
  geom_boxplot(fill = "lightpink", color = "black") +
  ggtitle("Overall Score by Mobile Phone Ownership") +
  xlab("Mobile Phone Ownership (0 = No, 1 = Yes)") +
  ylab("Overall Score")




# Box plot for OVERALL_SCORE vs UNIVERSITY
ggplot(df, aes(x = factor(UNIVERSITY), y = OVERALL_SCORE)) +
  geom_boxplot(fill = "lightcoral", color = "black") +
  ggtitle("Overall Score by University") +
  xlab("University") +
  ylab("Overall Score")


```


```{r}


t_test_computer <- t.test(OVERALL_SCORE ~ COMPUTER, data = df)
print(t_test_computer$p.value)

t_test_car <- t.test(OVERALL_SCORE ~ CAR, data = df)
print(t_test_car$p.value)

t_test_mobile <- t.test(OVERALL_SCORE ~ MOBILE, data = df)
print(t_test_mobile$p.value)

```



































